---
title: "Calibration Workflow with cuvis.r"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calibration Workflow with cuvis.r}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/CTTIR/cuvis.r/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/CTTIR/cuvis.r/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/CTTIR/cuvis.r/actions/workflows/pkgdown.yaml/badge.svg)](https://cttir.github.io/cuvis.r/)
[![CRAN status](https://www.r-pkg.org/badges/version/cuvis.r)](https://CRAN.R-project.org/package=cuvis.r)
[![Codecov test coverage](https://codecov.io/gh/CTTIR/cuvis.r/branch/main/graph/badge.svg)](https://app.codecov.io/gh/CTTIR/cuvis.r?branch=main)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/cuvis.r)](https://cran.r-project.org/package=cuvis.r)
[![CRAN downloads total](https://cranlogs.r-pkg.org/badges/grand-total/cuvis.r)](https://cran.r-project.org/package=cuvis.r)
[![License: Apache 2.0](https://img.shields.io/badge/License-Apache_2.0-blue.svg)](https://opensource.org/licenses/Apache-2.0)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

This vignette demonstrates the full calibration pipeline for Cubert
hyperspectral data: loading raw measurements, setting dark and white
references, and reprocessing to calibrated reflectance.

## Calibration Pipeline

Cubert cameras capture raw sensor data that must be calibrated through
several stages:
1. **Raw** -- unprocessed sensor readout
2. **Dark subtract** -- dark current noise removed
3. **Spectral radiance** -- calibrated to physical units (W/m^2/sr/nm)
4. **Reflectance** -- normalized by white reference (values 0--1)

## Step-by-Step Workflow

### 1. Initialize and Load Data

```{r init}
library(cuvis.r)

cuvis_init()

# Load the measurement session
session <- cuvis_session("path/to/measurement.cu3s")
mesu <- cuvis_get_measurement(session, 1)
```

### 2. Create Processing Context

The processing context holds calibration state and is created from the
session file. If the session already contains embedded references, they
are loaded automatically:

```{r ctx}
ctx <- cuvis_processing_context(session, load_references = TRUE)

# Check which references were loaded
cuvis_has_reference(ctx, "dark")
cuvis_has_reference(ctx, "white")
```

### 3. Set References (if needed)

If the session does not contain embedded references, load them from
separate files:

```{r refs}
# Load dark reference
dark_session <- cuvis_session("path/to/dark.cu3s")
dark_mesu <- cuvis_get_measurement(dark_session, 1)
cuvis_set_reference(ctx, dark_mesu, "dark")

# Load white reference
white_session <- cuvis_session("path/to/white.cu3s")
white_mesu <- cuvis_get_measurement(white_session, 1)
cuvis_set_reference(ctx, white_mesu, "white")
```

### 4. Reprocess to Reflectance

```{r reprocess}
cuvis_reprocess(ctx, mesu, mode = "reflectance")

# Extract the calibrated cube
cube <- cuvis_get_cube(mesu)
dim(cube)

# Reflectance values should be roughly in [0, 1]
range(cube, na.rm = TRUE)
```

### 5. Compare Processing Levels

```{r compare}
# Get raw data first
mesu_raw <- cuvis_get_measurement(session, 1)
cuvis_reprocess(ctx, mesu_raw, mode = "raw")
cube_raw <- cuvis_get_cube(mesu_raw)

# Compare a single pixel spectrum
pixel_raw <- cube_raw[50, 50, ]
pixel_ref <- cube[50, 50, ]
wavelengths <- attr(cube, "wavelengths")

plot(wavelengths, pixel_raw, type = "l", col = "gray",
     xlab = "Wavelength (nm)", ylab = "Value",
     main = "Raw vs Reflectance")
par(new = TRUE)
plot(wavelengths, pixel_ref, type = "l", col = "blue",
     axes = FALSE, xlab = "", ylab = "")
axis(4)
legend("topright", c("Raw", "Reflectance"),
       col = c("gray", "blue"), lty = 1)
```

### 6. Export Calibrated Data

```{r export}
# Export as ENVI (compatible with terra, stars, hsdar, etc.)
cuvis_export_envi(mesu, "output/calibrated/")

# Export as multi-channel TIFF
cuvis_export_tiff(mesu, "output/tiff/",
                  format = "MultiChannel", compression = "LZW")
```

### 7. Shutdown

```{r shutdown}
cuvis_shutdown()
```

## Integration with hyperspectR

If you have the [hyperspectR](https://github.com/r-heller/hyperspectR)
package installed, it uses cuvis.r as a backend:

```{r hyperspectr}
library(hyperspectR)

# Reads .cu3s files using cuvis.r internally
cube <- hs_read_cubert("measurement.cu3s")

# Full hyperspectR analysis pipeline
hs_plot_clinical(cube)
```
